Towards the near-infrared detection of brown dwarf companions: Exploring methods to detect low mass stellar companions from blended spectra \thanks{Based on observations collected at the European Organisation for Astronomical Research in the Southern Hemisphere under ESO programme 089.C-0977(A)}

### 2019-09-25 nir_companions.tex
Include my latest changes

The inverse survival function of the \(\chi^2\) distribution is used to determine the confidence levels on the minimum \(\chi^2\) parameters. The inverse survival function returns a \(\Delta\chi^2\) value from the minimum \(\chi^2\) value for a given sigma level and degree of freedom\footnote{In \textsc{python} with the \textsc{scipy} package this is a single line \texttt{scipy.stats.chi2(dof).isf(1-p)}, where \(p = 0.68\) for 1-\(\sigma\), and dof is the degree of freedom.}.
For example, the \(\Delta \chi^2\) for a single degree of freedom required for the 1-, 2-, and 3-\(\sigma\) levels is 1, 4, and 9 respectively~\citep{bevington_data_2003}. This method assumes that the measured flux is observed with a SNR sufficiently high so that the noise on the spectrum is approximately Gaussian, and the \(\chi^2\) method appropriate.

@@

We use the estimated companion temperatures from the Baraffe evolutionary models given the companion \(\rm M_2\) or \(\textrm{M}_2\sin{i}\) and stellar age as a starting point for the companion spectra grid computation and extend the grid typically by \pm\(500-1000\)~K in each direction, within the model limits based on trail and error. For example we show the companion temperature grid spanning \(-600\) to \(+400\)~K in Fig.~\ref{fig:Mdwarf_contours} and \(\pm400\)~K in Figs.~\ref{fig:HD211847_simulated_contours} and~\ref{fig:HD211847_result_contours}.

The large numbers stated above also do not include the RV grid for each component. The RV grid is user defined and the number of spectra /models to consider increases when the RV grid step size is decreased (finer RV resolution). We can reduce the RV grid space significantly by tailoring it to the target being examined. For each target, we use the estimated RV values from the observation time and orbital parameters, given in Table~\ref{tab:observations} as a centre starting point for the \({rv}_1\) and \({rv}_2\) values and increment the RV within a few FWHM around those values, or out to the targets \(\rm K_1\) and estimated \(\rm K_2\) values.

@@

For {HD 30501} the injection onto synthetic and observed spectra produce similar results. At temperatures above 3800~K in both the real and synthetic the injected companion is recovered within 100~K. For injected companion temperatures below 3800~K the temperature recovered is systematically higher than the injected value. This indicates that the companion is not correctly recovered and is affected by the added noise. We determine this temperature to be the upper temperature limit for the recovery. For the other stars we could not conclude on the upper limit due to spectral mismatch issues. In these cases we use the results from the synthetic injection to derive a temperature recovery cut-off for each target, each simulated with the closest host star spectrum.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\hsize]{images/fig8.pdf}
    \caption{(top) Companion temperature verses \(\chi^2\) for simulations with different injected companion temperatures. Other fixed parameters for these fully synthetic simulations was \(T_{\textrm{eff}_1}=5200\)~K, \(\textrm{logg}_1=4.5\), \(\textrm{logg}_2=5.0\), and both [Fe/H] = 0.0. A fixed Gaussian noise corresponding to a SNR of 300 was used.
        (bottom) A close up view of \(\chi^2\) < 15. The three horizontal grey lines indicate the 1, 2, 3 sigma with 2 degrees of freedom. The vertical dotted lines indicate the location of the minimum \(\chi^2\) recovered for each companion. The black solid vertical in both panels shows the 2300~K cut-off of the PHOENIX-ACES models}
    \label{fig:injection_shape}
\end{figure}

@@

Using the temperature cut-off values, we derive an upper mass limit for the companions around our stars using the~\citet{baraffe_new_2015} evolutionary models, finding the closest point matching the spectral temperature cut-off and \(\rm logg=5.0\). These values are given in Table~\ref{tab:mass_limits} and are between 560 and 618 \(\rm M_{Jup}\). The flux ratio between the cut-off and the host star are also provided for, being between 5 and 15 per cent in this wavelength span.

@@

\subsubsection{Line contribution of faint companions}
\label{subsubsec:line_contributions}
We calculate the line depths of the synthetic companion spectra to determine the SNR levels required to detect the lines of the binary companions.
One thing easy to overlook when attempting to detect the binary companion at low flux ratios is the actual contribution of the spectral lines of the companion.
The flux ratio of the continuum for our most promising target is \(\rm F_2/F_1\)\(\sim\)1 per cent with the other targets having an expected flux ratio 0.5 per cent, and some well below. The spectral lines of the individual components which are the features we are trying to detect with the binary model, have depths on average around 10--20 per cent of their respective continua,  at-least between 2110--2160~nm. In effect, the companion line features have a depth \(\ll\) 1 per cent relative to the continuum of the combined spectra.

In Table~\ref{tab:line_contributions} we calculate some properties of the spectral lines in the PHOENIX-ACES library between 2110--2160~nm. We count the number of spectral lines (\emph{no.\ lines}) deeper than 5 per cent, and take the average depth (\emph{avg.\ depth}) of these lines. The contribution \emph{cont.\ depth} of the companion lines to a combined spectrum accounts for the flux ratio between the two components. Here we use a Sun-like host with \(T_{\textrm{eff}_1}=5800\)~K. This simplified combination neglects the continuum shapes of both spectral components and uses the average flux ratios for this wavelength range. The PHOENIX-ACES spectra in the temperature range of 2500--5000~K shown in Fig.~\ref{fig:comp_spectra} can be used to get a visual indication of the line density and depth measured here.

@@

The SNR of the observed spectra is between 150--350, which is below the SNR of 323 needed for the detection of the low-mass star companion of {HD 211847} with temperature 3200~K and logg 5.0. For our other targets with BD companions at and below the PHOENIX-ACES temperature range, we would need observed SNR >800 to detect the individual spectral lines of the companion. With the SNR increasing with \(\sqrt{N}\) this would require the observational time for each target to be increased by a factor of \(\sim\)10--64.
{This is in line with the recent detection of the spectrum of a non-transiting giant planet by~\citet{piskorz_evidence_2016} which utilized nIR spectra with SNR > 2000, from 1--3 hours of observation each.}

>>>>

The inverse survival function of the \(\chi^2\) distribution is used to determine the confidence levels on the minimum \(\chi^2\) parameters. The inverse survival function returns a \(\Delta\chi^2\) value from the minimum \(\chi^2\) value for a given $sigma$-level and degree of freedom\footnote{In \textsc{python} with the \textsc{scipy} package this is a single line \texttt{scipy.stats.chi2(dof).isf(1-p)}, where \(p = 0.68\) for 1-\(\sigma\), and dof is the degree of freedom.}.
For example, the \(\Delta \chi^2\) for a single degree of freedom required for the 1-, 2-, and 3-\(\sigma\) levels is 1, 4, and 9 respectively~\citep{bevington_data_2003}. This method assumes that the measured flux is observed with a SNR sufficiently high so that the noise on the spectrum is approximately Gaussian, and the \(\chi^2\) method appropriate.

@@

We use the estimated companion temperatures from the Baraffe evolutionary models given the companion \(\rm M_2\) or \(\textrm{M}_2\sin{i}\) and stellar age as a starting point for the companion spectra grid computation and extend the grid typically by \(\pm500-1000\)~K in each direction, within the model limits based on trail and error. For example we show the companion temperature grid spanning \(-600\) to \(+400\)~K in Fig.~\ref{fig:Mdwarf_contours} and \(\pm400\)~K in Figs.~\ref{fig:HD211847_simulated_contours} and~\ref{fig:HD211847_result_contours}.

The large numbers stated above also do not include the RV grid for each component. The RV grid is user defined and the number of spectra/models to consider increases when the RV grid step size is decreased (finer RV resolution). We can reduce the RV grid space significantly by tailoring it to the target being examined. For each target, we use the estimated RV values from the observation time and orbital parameters, given in Table~\ref{tab:observations} as a centre starting point for the \({rv}_1\) and \({rv}_2\) values and increment the RV within a few FWHM around those values, or out to the targets \(\rm K_1\) and estimated \(\rm K_2\) values.

@@

For {HD 30501} the injection onto synthetic and observed spectra produce similar results. At temperatures above 3800~K in both the real and synthetic the injected companion is recovered within 100~K. For injected companion temperatures below 3800~K the temperature recovered is systematically higher than the injected value. This indicates that the companion is not correctly recovered and is affected by the added noise. We determine this temperature to be the upper temperature limit for the recovery. For the other stars in this work we could not conclude on the upper limit due to spectral mismatch issues. In these cases we use the results from the synthetic injection to derive a temperature recovery cut-off for each target, each simulated with the closest host star spectrum.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\hsize]{images/fig8.pdf}
    \caption{(top) Companion temperature verses \(\chi^2\) for simulations with different injected companion temperatures. Other fixed parameters for these fully synthetic simulations was \(T_{\textrm{eff}_1}=5200\)~K, \(\textrm{logg}_1=4.5\), \(\textrm{logg}_2=5.0\), and both [Fe/H] = 0.0. A fixed Gaussian noise corresponding to a SNR of 300 was used.
    (bottom) A close up view of \(\chi^2\) < 15. The three horizontal grey lines indicate the 1-, 2-, 3-$\sigma$ for 2 degrees of freedom. The vertical dotted lines indicate the location of the minimum \(\chi^2\) recovered for each companion. The black solid vertical in both panels shows the 2300~K cut-off of the PHOENIX-ACES models}
    \label{fig:injection_shape}
\end{figure}

@@

Using the temperature cut-off values, we derive an upper mass limit for the companions around our stars using the~\citet{baraffe_new_2015} evolutionary models, finding the closest point matching the spectral temperature cut-off and \(\rm logg=5.0\). These values are given in Table~\ref{tab:mass_limits} and are between 560 and 618 \(\rm M_{Jup}\). The flux ratio between the cut-off temperature and the host star was also calculated to be between 5 and 15 per cent in this wavelength range.

@@

\subsubsection{Line contribution of faint companions}
\label{subsubsec:line_contributions}
We preformed calculations of the actual line depths of the companion in the synthetic binary spectrum to determine the SNR levels required to directly detect the lines of the binary companions.
One thing easy to overlook when attempting to detect the binary companion at low flux ratios is the actual contribution of the spectral lines of the companion.
The flux ratio of the continuum for our most promising target is \(\rm F_2/F_1\)\(\sim\)1 per cent with other targets having an expected flux ratio 0.5 per cent, and some well below. The spectral lines of the individual components, which are the features we are trying to detect, have depths on average around 10--20 per cent of their respective continua, at-least between 2110--2160~nm. In effect, the companion line features have a depth \(\ll\) 1 per cent relative to the continuum of the combined spectra.

In Table~\ref{tab:line_contributions} we calculate some properties of the spectral lines in the PHOENIX-ACES library between 2110--2160~nm. We count the number of spectral lines (\emph{no.\ lines}) deeper than 5 per cent, and take the average depth (\emph{avg.\ depth}) of these lines. The contribution \emph{comp.\ depth} of the companion lines to a combined spectrum accounts for the flux ratio between the two components. Here we use a Sun-like host with \(T_{\textrm{eff}_1}=5800\)~K. This simplified combination neglects the continuum shapes of both spectral components and uses the average flux ratios for this wavelength range. The PHOENIX-ACES spectra in the temperature range of 2500--5000~K shown in Fig.~\ref{fig:comp_spectra} can be used to get a visual indication of the line density and depths used here.

@@

The SNR of the observed spectra here is around 150--350, which is below the SNR of 323 needed to directly observe the spectral lines of the low-mass companion with temperature 3200~K and logg 5.0 in the spectrum of {HD 211847} at a 1-$\sigma$ level. For our other targets with BD companions at and below the PHOENIX-ACES temperature range, we would need observed SNR >800 to directly detect the individual spectral lines of the companion. With the SNR increasing with \(\sqrt{N}\sim\sqrt{t_{exp}}\) this would require the observational time for each target to be increased by a factor of \(\sim\)10--64.
{This is in line with the recent detection of the spectrum of a non-transiting giant planet by~\citet{piskorz_evidence_2016} which utilized nIR spectra with SNR > 2000 resulting from 1--3 hours of exposure time each.}
